---
title: "new_comparing_cluster"
author: "Veronica Mandelli"
date: "4/28/2020"
output: html_document
---


To run this code some self made functions need to be loaded
```{r pressure, results='hide',warning=FALSE, message=FALSE}
#load libraries
library(easypackages)
library(fpc)
libraries("here","ggplot2","psych","gplots","ramify","cluster","readxl","pracma","MatchIt","grid","amap","gridExtra",
          "reticulate","plsgenomics","clusteval",'gridExtra')

# set paths
codepath = here("code")
tidydatapath = here("data","tidy")
rowdatapath = here("data")
subjectiddatapath = here ("data","SUBJECTID")
plotpath = here("plots")
resultpath = here("results","FINAL")
code_nosharepath = here('code_noshare')
```

### load DBs
and make varibales  homogeneous
```{r, message=FALSE, warning= FALSE, results= 'hide', include=FALSE}
# NDAR
# read the .csv file
DB_NDAR = read.csv(file.path(tidydatapath,"NDAR_cleaned_withNaN.csv"),header = TRUE, stringsAsFactors = FALSE)
DB_NDAR_demo = DB_NDAR #save other varibles for late analysis

# select only needed columns
colnames = c("subjectkey","target","interview_age","sex","mabc2_manualdexterity_std",
             "mabc2_aimingcatching_std","mabc2_balance_std","mabc2_total_std",'title')
DB_NDAR = DB_NDAR[,colnames]

# change variables names to make easier to compare DBs
colnames(DB_NDAR)[which(names(DB_NDAR) == "subjectkey")] <- "subject_id"
colnames(DB_NDAR)[which(names(DB_NDAR) == "interview_age")] <- "age"
colnames(DB_NDAR)[which(names(DB_NDAR) == "title")] <- "study_id"

# transform age in years
DB_NDAR$age=round(DB_NDAR$age/12,2)

# change study id code to make it more readable
for (i in (1:144)) {
    if (DB_NDAR[i, 'study_id'] == "Direct Examination of Imitation-Based Learning in Autism"){
      DB_NDAR[i,'study_id'] = 'NDAR_DEML'
    }
    else if (DB_NDAR[i, 'study_id'] =="A Novel Framework for Impaired Imitation in ASD" ){
      DB_NDAR[i,'study_id'] = 'NDAR_NFII'
    }
    else if (DB_NDAR[i, 'study_id'] == "The neurobiological basis of heterogeneous social and motor deficits in ASD"){
      DB_NDAR[i,'study_id'] = 'NDAR_NBSM'}
}

# exclude people older than 16 years old
DB_NDAR = DB_NDAR[DB_NDAR$age< 17,] # 3 subjects

# BP

DB_BP =  read_excel(file.path(tidydatapath,"BP_abc.xlsx"),
    col_types = c("text", "text","text","numeric", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric" ))

DB_BP_demo = DB_BP #save other varibles for late analysis

# select only needed columns

colnames = c("ID","diagnosis",'age_years','sex',"MABC2_ManualDexterity_standardscores",
             "MABC2_AimingCatching_standardscores","MABC2_Balance_standardscores","MABC2_Total_StandardScores")
DB_BP = DB_BP[,colnames]


# mask for na in ASD group before clustering
# cols2use  = c("abc_child_manualdexterity","abc_child_pointthrow","abc_child_balance")
cols2use = c("MABC2_ManualDexterity_standardscores","MABC2_AimingCatching_standardscores","MABC2_Balance_standardscores")
mask = rowSums(is.na(DB_BP[,cols2use]))==0
DB_BP = as.data.frame(DB_BP[mask,])

# change variables names to make easier to compare DBs
# colnames(DB_BP)[which(names(DB_BP) == "group")] <- "target"
colnames(DB_BP)[which(names(DB_BP) == "MABC2_ManualDexterity_standardscores")] <- "mabc2_manualdexterity_std"
colnames(DB_BP)[which(names(DB_BP) == "MABC2_AimingCatching_standardscores")] <- "mabc2_aimingcatching_std"
colnames(DB_BP)[which(names(DB_BP) == "MABC2_Balance_standardscores")] <- "mabc2_balance_std"
colnames(DB_BP)[which(names(DB_BP) == "MABC2_Total_StandardScores")] <- "mabc2_total_std"
colnames(DB_BP)[which(names(DB_BP) == "ID")] <- "subject_id"
colnames(DB_BP)[which(names(DB_BP) == "age_years")] <- "age"
colnames(DB_BP)[which(names(DB_BP) == "diagnosis")] <- "target"

# add study id code
DB_BP$study_id = 'BP'

# make sex homogenous  across DBs
for (k in (1:length(DB_BP$sex))) {
   if (DB_BP[k,'sex'] == "0"){
       DB_BP[k,'sex'] = 'M'}
    else {DB_BP[k,'sex'] = "F"}
}

# select only children older than 3 y.o  (ABC-movement not standardized for younger!)
DB_BP = DB_BP[DB_BP$age>=3,] # 3 subject
```

### modules inferring
The ABC movement battery is dividend into 3 modules selected based on the age of the child. Since the each module has specific tasks here we would like to correct for possible differences due to this difference
module 1: 3-6 years old
module 2: 7-10 years old
module 3: 11-16 years old
```{r,results='hide', warning= FALSE, message=FALSE}

# NDAR
data=DB_NDAR

  for (i in (1:length(data$subject_id))) {
    if (data[i, 'age']<7){
      data[i,'module'] = 1
    }
    else if (data[i, 'age']>=7 && data[i, 'age'] < 11){
      data[i,'module'] = 2
    }
    else if (data[i, 'age']>=11){
      data[i,'module'] = 3}
    
  }

DB_NDAR=data

# BP
data=DB_BP


  for (i in (1:length(data$subject_id))) {
    if (data[i, 'age']<7){
      data[i,'module'] = 1
    }
    else if (data[i, 'age']>=7 && data[i, 'age'] < 11){
      data[i,'module'] = 2
    }
    else if (data[i, 'age']>=11){
      data[i,'module'] = 3}
    
  }

DB_BP=data

```

### merge DBs in one big dataset
```{r,results='hide', warning= FALSE, message=FALSE}
# merge the DBs
cols2use = c ('subject_id','module','sex','target','study_id','age','mabc2_manualdexterity_std','mabc2_aimingcatching_std',
              'mabc2_balance_std','mabc2_total_std')

dataset = rbind(DB_BP[,cols2use],DB_NDAR[,cols2use])
# reset row names
rownames(dataset) <- 1:nrow(dataset)

#save sub-ids for NDA
subject_list  = dataset[dataset$study_id %in% c("NDAR_DEML","NDAR_NFII","NDAR_NBSM"), c("subject_id","study_id","target")]
library('openxlsx')
#write.xlsx(subject_list,file.path(resultpath,'NDA_subids.xlsx'))
```

### beta correction
apply a beta correction to remove the variance in the data given by different recruiting sites,sex and module of the ABC used.
```{r,results='hide', warning= FALSE, message=FALSE}

# effect of interest : ABC scores 

cols2use = c ('mabc2_manualdexterity_std','mabc2_aimingcatching_std','mabc2_balance_std','mabc2_total_std')

# rename columns and change format
data2use = dataset[,c('module','sex','target','study_id','age', cols2use)]
data2use$sex =  as.factor(data2use$sex)
data2use$module =  as.factor(data2use$module)
data2use$target =  as.factor(data2use$target)
data2use$study_id =  as.factor(data2use$study_id)

# for each column (3 ABC score + total ABC)
for (i in c(1:length(cols2use))){
  y_var = cols2use[i]
  lm_formula = as.formula(sprintf( "%s~ %s + %s + %s + %s", y_var ,'target','module','sex','study_id'))
  fit  =  eval(substitute(lm(formula= lm_formula, data = data2use)))

  covname2use = c('module2', 'module3','sexM','study_idNDAR_DEML','study_idNDAR_NBSM','study_idNDAR_NFII')         
  beta1 = fit$coefficients[covname2use, drop = FALSE]
  beta1[is.na(beta1)] = 0
  
  full_model =  model.matrix(~0 + as.factor (module) 
                           + as.factor (sex)
                           + as.factor (study_id),
                           data = data2use)
  colnames(full_model) = c('module1','module2', 'module3','sexM','study_idNDAR_DEML','study_idNDAR_NBSM','study_idNDAR_NFII') 
  col_name = paste('new_',y_var,sep='')
  
  data2use[,col_name] =  as.numeric(t(data2use[,y_var]) - beta1 %*% t(full_model[,covname2use]))
  
}

dataset_corr = data2use [,c("target",'study_id','age','sex','module',"new_mabc2_manualdexterity_std",
                            "new_mabc2_aimingcatching_std","new_mabc2_balance_std" ,
                          'new_mabc2_total_std')]
```

### re change of columns names
```{r,results='hide', warning= FALSE, message=FALSE}
dataset_corr$subject_id = dataset$subject_id
dataset_corr$age = dataset$age

colnames(dataset_corr)[which(names(dataset_corr) == "new_mabc2_manualdexterity_std")] <- "mabc2_manualdexterity_std"
colnames(dataset_corr)[which(names(dataset_corr) == "new_mabc2_aimingcatching_std")] <- "mabc2_aimingcatching_std"
colnames(dataset_corr)[which(names(dataset_corr) == "new_mabc2_balance_std")] <- "mabc2_balance_std"
colnames(dataset_corr)[which(names(dataset_corr) == "new_mabc2_total_std")] <- "mabc2_total_std"
```

### selecting only ASD to clustering and saving
```{r,results='hide', warning= FALSE, message=FALSE}
dataset_asd= subset(dataset_corr, dataset_corr$target=='ASD')
#save all corrected dataset
col2save = c("subject_id" ,"sex" ,"study_id","module","mabc2_manualdexterity_std", "mabc2_aimingcatching_std", "mabc2_balance_std", "mabc2_total_std")
savingpath_reval="/Users/vmandelli/Google Drive/PHD/projects/motorNDAR/results/REVAL"
#write.csv(dataset_asd[,col2save],file.path(savingpath_reval,"data4reval.csv"))
write.csv(dataset_corr,file.path(resultpath,"dataset_complete_corrected.csv"))
```

## END