---
title: "R Notebook"
output: html_notebook
author: Veronica Mandelli
date: 18.03.2021
---
In this script we produce visualization (and statistics) for describing the ASD motor clusters comparing them to the TD group

```{r pressure, results='hide',warning=FALSE, message=FALSE}
# load librayes
library(easypackages)
#library(fpc)
libraries("here","ggplot2","psych","gplots",'car',"ramify","NbClust","cluster","readxl","plot3D","beeswarm",'gridExtra',"ggplot",'openxlsx')
# set paths
codepath = here("code")
subjectiddatapath = here ("data","SUBJECTID")
plotpath = here("plots","FINAL")
resultpath = here("results","FINAL")
code_nosharepath = here('code_noshare')
reval_result_data_path = here("results","reval_90_server")
tidydatapath = here("data","tidy")

#useful fucntions
source(file.path(code_nosharepath,"cohens_d.R"))
source(file.path(code_nosharepath,"Plot_3d.R"))
source(file.path(code_nosharepath,"Gplot.R"))
source(file.path(code_nosharepath,"other_features.R"))
source(file.path(code_nosharepath,"new_column_4clust.R"))


color = c("#FF9933","#3399CC","#009900","#ff0000","#000000")
```

### import clusters label from reval + TD dataset
```{r, warning=FALSE, message=FALSE}
#import labels
trts_dataset = read.csv(file.path (reval_result_data_path,'reval_results_90.csv'),header = TRUE, stringsAsFactors = FALSE)
#change cols names
colnames(trts_dataset)[which(names(trts_dataset) == "subjectkey")] <- "subject_id"

#merge labels with variables DB
dataset_complete_asd_td = read.csv(file.path(tidydatapath,"dataset_complete_asd_td_ndar_bp.csv"),header = TRUE, stringsAsFactors = FALSE)  

# merge the data with clusters ASD only
merged_dataset_ASD = merge(trts_dataset, subset(dataset_complete_asd_td,dataset_complete_asd_td$target=='ASD') ,by = "subject_id")

# ADD TD
dataset_td = subset(dataset_complete_asd_td,dataset_complete_asd_td$target=='TD') 
dataset_td$cluster = "TD"
dataset_td$TR_TS = "TD"
merged_dataset_ASD_TD = rbind(merged_dataset_ASD,dataset_td)
```

# manually change name to the clusters
```{r}
clust2use = "cluster"
# subset data to learn cluster order from the train
data2use_tr= subset(merged_dataset_ASD_TD,merged_dataset_ASD_TD$TR_TS=="TR")
data_tr = data2use_tr
# learn the order  from the train (they are homolougous (as python default))
avg_tot = (tapply(data2use_tr$mabc2_manualdexterity_std, data2use_tr[,clust2use], mean))
group_order = rownames(as.data.frame(sort(avg_tot)))
#print(group_order)

#change the order in train
data_2b_corrected = subset(merged_dataset_ASD_TD,merged_dataset_ASD_TD$TR_TS %in% c("TS","TR"))
  for (j in (1:nrow(data_2b_corrected))) {
    if (data_2b_corrected[j, clust2use]== group_order[1]){
      data_2b_corrected[j,paste("new_",clust2use,sep="")] = "low"
    }
    else if (data_2b_corrected[j, clust2use]==group_order[2]){
      data_2b_corrected[j,paste("new_",clust2use,sep="")] = "high"
    }
}

dataset_corr = data_2b_corrected

# add the DB
dataset_td = subset(merged_dataset_ASD_TD,merged_dataset_ASD_TD$target=='TD') 
dataset_td[,paste("new_",clust2use,sep="")] = "TD"
complete_dataset_corr = rbind(dataset_corr,dataset_td)

```

### plot cluster description
```{r}
source(file.path(code_nosharepath,"Gplot.R"))
new_clust2use = "new_cluster" 
what_group = "TR"
color = c( "#72aeb6","#2f70a1","#f0be3d")
data2use= subset(dataset_corr,dataset_corr$TR_TS==what_group)
tr_plot = Gplot(data2use,color,new_clust2use,what_group)

what_group = "TS"
data2use= subset(dataset_corr,dataset_corr$TR_TS==what_group)
ts_plot = Gplot(data2use,color,new_clust2use,what_group)

g =  grid.arrange(tr_plot[[1]],ts_plot[[1]],tr_plot[[2]],ts_plot[[2]],tr_plot[[3]],ts_plot[[3]],tr_plot[[4]],ts_plot[[4]], nrow = 4)
#ggsave(file.path(plotpath, "cluster_description_TR_TS_vertical.png"), g,width = 4,height = 8)
```

### statistics to describe MABC2 features in the clusters
```{r}
#TRAIN
#cohens'd
data2use= subset(dataset_corr,dataset_corr$TR_TS=='TR')
cohens_d_mat = as.data.frame(matrix(ncol=1,nrow=4))
rownames(cohens_d_mat) = c('mabc2_manualdexterity_std','mabc2_aimingcatching_std', 'mabc2_balance_std', 'mabc2_total_std')
for (col in c('mabc2_manualdexterity_std','mabc2_aimingcatching_std', 'mabc2_balance_std', 'mabc2_total_std')){
CohenD <- cohens_d(data2use[data2use$new_cluster=='high',col],
                       data2use[data2use$new_cluster=='low',col], 
                       SIGN=FALSE)
   
      cohens_d_mat[col,1] = CohenD}

#write.csv(cohens_d_mat, file.path(resultpath,'cohensd_cluster_TR.csv'))

#TEST
#cohens'd
data2use= subset(dataset_corr,dataset_corr$TR_TS=='TS')
cohens_d_mat = as.data.frame(matrix(ncol=1,nrow=4))
rownames(cohens_d_mat) = c('mabc2_manualdexterity_std','mabc2_aimingcatching_std', 'mabc2_balance_std', 'mabc2_total_std')
for (col in c('mabc2_manualdexterity_std','mabc2_aimingcatching_std', 'mabc2_balance_std', 'mabc2_total_std')){
CohenD <- cohens_d(data2use[data2use$new_cluster=='high',col],
                       data2use[data2use$new_cluster=='low',col], 
                       SIGN=FALSE)
   
      cohens_d_mat[col,1] = CohenD}

#write.csv(cohens_d_mat, file.path(resultpath,'cohensd_cluster_TS.csv'))

# Percentage table of clusters number
tabperc = table(dataset_corr$new_cluster)/nrow(dataset_corr)*100
#write.csv(tabperc, file.path(resultpath,'clust_perc.csv'))
```


### load the IRCCS-MEDEA demographic dataset, which contains non-motor characteristics of the subject. Comparre them in the clusters
```{r}
# load BP demo data to get info about other features and merged it with the clusters
BP_demo = read.csv(file.path (tidydatapath,'DB_BP_demo_renamed.csv'),header = TRUE, stringsAsFactors = FALSE)
col2use=c("age_years","IQ" ,"SRS_TotalTScores","ADOS_TotalScore","ADOS_SocialAffect", "ADOS_RestrictedRepetitiveBehavior",'ID', "SRS_SocialAwareness","SRS_SocialCognition","SRS_SocialCommunication","SRS_SocialMotivation","SRS_RRB")
title_list = c('AGE','IQ','SRS','ADOS CSS TOTAL','ADOS CSS SA', 'ADOS CSS RRB', "SRS_SocialAwareness","SRS_SocialCognitio","SRS_SocialCommunication","SRS_SocialMotivation","SRS_RRB","early motor dev","early lang dev")          
data2plot_ = merge(dataset_corr[dataset_corr$study_id=="BP",],BP_demo[,col2use],by.x ="subject_id",by.y="ID")
color2use = c("#FF9933","#3399CC","#009900","#ff0000","#000000")

# load database with information about the early development
early_dev_BP = read_excel(file.path(tidydatapath,"SupplementaryData_EarlyMotor-Language.xlsx"))

# change subjectid format 
early_dev_BP$ID <- gsub("RFASD_", "1",early_dev_BP$ID)
early_dev_BP$ID <- gsub("RFasd_", "1",early_dev_BP$ID)
# select columns
early_dev_BP=early_dev_BP[,c("ID","mot_early","lang_early")]

# merge 
data2plot = merge(early_dev_BP,data2plot_, by.x="ID",by.y = 'subject_id')
#var_list = c("mot_early","lang_early")


# define columns to plot 
cols2plot =  c("age_years","IQ" ,"SRS_TotalTScores","ADOS_TotalScore","ADOS_SocialAffect", "ADOS_RestrictedRepetitiveBehavior","SRS_SocialAwareness","SRS_SocialCognition","SRS_SocialCommunication","SRS_SocialMotivation","SRS_RRB","mot_early","lang_early")
p_list=c()

# make columns in the format required to be plotted
data2plot$new_cluster<- factor( data2plot$new_cluster,levels = c('high','low'));
data2plot$sex <- factor( data2plot$sex,levels = c("F","M"))

# prepare table to save statistic
other_feat_mat= as.data.frame(matrix(nrow=length(cols2plot),ncol = 7))
rownames(other_feat_mat) = cols2plot
colnames(other_feat_mat) = c("normality","levene","test_type","t","p","df","CohenD")

# run plots and stats in loop
for (i in 1:length(cols2plot)){
  
  var2use = cols2plot[i]
  print(var2use)
  print(sum(!is.na(data2plot[,var2use])))
  #plot
  p = ggplot(data =data2plot ,aes_string(x = 'new_cluster', y = var2use, color='new_cluster') )
  p = p + geom_jitter() + geom_boxplot(fill = NA, colour = "#000000", outlier.shape = NA)
  p = p + scale_colour_manual(values = color)
  p = p + guides(colour = FALSE)
  p = p + xlab("") + ylab('')
  p = p + ggtitle(title_list[i])
  p = p + theme(plot.title = element_text(size = 7, face = "bold",hjust = 0.5))
  p = p + theme(axis.text.x = element_text(face="bold", size=7))
  p = p + theme(axis.text.y = element_text(face="bold", size=7))
  
  print(p)
  #ggsave(file.path(plotpath, paste("othfeat_",var2use,".png",sep='')), p,width = 2,height = 2)
  p_list[[i]]= p
  
  #statistics
  # t-test requirments
  
  # Compute Shapiro wilk test by goups --> normality
  resp_normality = shapiro.test(data2plot[,var2use])
  #resp_normality
  other_feat_mat[var2use,"normality"] = resp_normality$p.value
  
  # Equality of variance
  
  # Levene's test with one independent variable
  df2<-data2plot[complete.cases(data2plot[,c(var2use,'new_cluster')]),]
  form2use = as.formula(sprintf("%s ~ new_cluster",var2use))
  
  resp_levene = leveneTest(form2use, data = df2)
  #resp_levene
  other_feat_mat[var2use,"levene"] = resp_levene$`Pr(>F)`[1]
  
  if (  resp_normality$p.value >0.05 && resp_levene$`Pr(>F)`>0.05 ){
    #  paramentric: t-test
    x = subset(data2plot, data2plot$"new_cluster" =="high")[,var2use]
    y = subset(data2plot, data2plot$"new_cluster" =="low")[,var2use]
    res_t = t.test(x,y)
    #print(res_t)
    other_feat_mat[var2use,"test_type"] = 't-test'
    other_feat_mat[var2use,"t"] = res_t$statistic
    other_feat_mat[var2use,"p"] = res_t$p.value
    other_feat_mat[var2use,"df"] = res_t$parameter
  }
  else{# non paramentric:
    x = subset(data2plot, data2plot$"new_cluster" =="high")[,var2use]
    y = subset(data2plot, data2plot$"new_cluster" =="low")[,var2use]
    res_w = wilcox.test(x, y, alternative = "two.sided")
    
    other_feat_mat[var2use,"test_type"] = 'wilcox'
    other_feat_mat[var2use,"t"] = res_w$statistic
    other_feat_mat[var2use,"p"] = res_w$p.value
  }
  
  
  # we also compute the cohens'd to estimate the effect size
  CohenD <- cohens_d(data2plot[data2plot$new_cluster=='high',var2use],
                     data2plot[data2plot$new_cluster=='low',var2use], 
                     SIGN=FALSE)
  
  other_feat_mat[var2use,"CohenD"] = CohenD
  
}


# all plots in a unique figure
new_grid = grid.arrange(
  p_list[[1]],p_list[[2]],p_list[[3]],p_list[[4]],p_list[[5]],p_list[[6]],p_list[[12]],p_list[[13]],
  nrow = 2)
#ggsave(file.path(plotpath,"other_feat_BP.png"),new_grid,width = 6,height = 4)

# save statistics
other_feat_mat$variable = rownames(other_feat_mat)
#write.xlsx(other_feat_mat,file.path(resultpath,'FORCED_other_features_ttest_1102023.xlsx'))
```

### save IRCCS-MEDEA dataset with clusters labels for the following kinematic analysis
```{r}
data_BP = subset(dataset_corr,dataset_corr$study_id=="BP")
#savingpath_dataset=""
write.csv(data_BP,file.path(resultpath,"dataset_bp_ASD_TD_clusters.csv"))
```

### statistics comparing Autism Subtypes (High and Low) with TD and DCD grups
```{r}
#require(tidyverse)
library(rstatix)
library(ggpubr)
library(car)
alldatset = read.csv(file.path(resultpath,"dataset_complete_corrected.csv"),header = TRUE, stringsAsFactors = FALSE)
alldatset = merge(alldatset, trts_dataset,by= 'subject_id',all.x = TRUE,all.y=TRUE)
alldatset[alldatset$target=="TD",'cluster']="TD"
alldatset[alldatset$target=="DCD",'cluster']="DCD"
alldatset[alldatset$cluster==0,'cluster']="high"
alldatset[alldatset$cluster==1,'cluster']="low"


#anova
# Compute the analysis of variance
res.aov <- aov(mabc2_total_std ~ cluster, data = alldatset)
# Summary of the analysis
summary(res.aov)
AIC(res.aov)


col2use =  c("mabc2_total_std" )
all_sub_clust = alldatset
# prepare table to save statistic
other_feat_mat= as.data.frame(matrix(nrow=4,ncol = 7))
rownames(other_feat_mat) = c("high_TD","high_DCD","low_TD","low_DCD")
colnames(other_feat_mat) = c("normality","levene","test_type","t","p","df","CohenD")
list_comarison_a = c("high","high","low","low")
list_comarison_b =c("TD","DCD","TD","DCD")
row_list = c("high_TD","high_DCD","low_TD","low_DCD")
# t-test requirments
for (i in 1:4 ){
row2use = row_list[i]
comp2use_a = list_comarison_a[i]
comp2use_b = list_comarison_b[i]

# t-test requirments
      all_sub_clust_subset =  subset(all_sub_clust,all_sub_clust$cluster %in% c(comp2use_a,comp2use_b))
      # Compute Shapiro wilk test by goups --> normality
      resp_normality = shapiro.test(all_sub_clust_subset[,col2use])
      #resp_normality
      other_feat_mat[row2use,"normality"] = resp_normality$p.value
      
      # Equality of variance
      
      # Levene's test with one independent variable
      form2use = as.formula(sprintf("%s ~ cluster",col2use))
      resp_levene = leveneTest(form2use, data = all_sub_clust_subset)
      #resp_levene
      other_feat_mat[row2use,"levene"] = resp_levene$`Pr(>F)`[1]
      
      if (resp_normality$p.value >0.05 && resp_levene$`Pr(>F)`>0.05 ){
      #  paramentric: t-test
      x = subset(all_sub_clust, all_sub_clust$"cluster" ==comp2use_a)[,col2use]
      y = subset(all_sub_clust, all_sub_clust$"cluster" ==comp2use_b)[,col2use]
      res_t = t.test(x,y)
      #print(res_t)
      other_feat_mat[row2use,"test_type"] = 't-test'
      other_feat_mat[row2use,"t"] = res_t$statistic
      other_feat_mat[row2use,"p"] = res_t$p.value
      other_feat_mat[row2use,"df"] = res_t$parameter
      }
      else{# non paramentric:
      x = subset(all_sub_clust, all_sub_clust$"cluster" ==comp2use_a)[,col2use]
      y = subset(all_sub_clust, all_sub_clust$"cluster" ==comp2use_b)[,col2use]
      res_w = wilcox.test(x, y, alternative = "two.sided")

      print(res_t)
      other_feat_mat[row2use,"test_type"] = 'wilcox'
      other_feat_mat[row2use,"t"] = res_w$statistic
      other_feat_mat[row2use,"p"] = res_w$p.value}

      
      # we also compute the cohens'd to estimate the effect size
      CohenD <- cohens_d(all_sub_clust[all_sub_clust$cluster==comp2use_a,col2use],
                       all_sub_clust[all_sub_clust$cluster==comp2use_b,col2use], 
                       SIGN=FALSE)
      
      other_feat_mat[row2use,"CohenD"] = CohenD

}
other_feat_mat


write.csv(other_feat_mat,file.path(resultpath,'ttest_cluster_TDDCD.csv'))
```

# reval stability plot - plot it in a nicer formtat
```{r}
reval = read.csv(file.path(reval_result_data_path,"metric_90.csv"))
reval_val = reval[,c('cl','stab_val',"stab_val_low")]
colnames(reval_val)=c("k","value","CI")
reval_val$cat = 'validation'

reval_tr = reval[,c('cl','stab_tr',"stab_tr_low")]
colnames(reval_tr)=c("k","value","CI")
reval_tr$cat = 'train'
mat = rbind(reval_tr,reval_val)
colors = c("#cc0000","#000099")
# spaghetti plot
gg.spaghetti_x<-ggplot(mat, aes (x = k, y =value , group = cat, color = cat))
gg.spaghetti_x <-gg.spaghetti_x +geom_line() + geom_point()   +
                                 geom_errorbar(aes(ymin= value - CI, ymax= value + CI), width = 0.1, size = 0.5) + 
                                 theme(axis.text.x = element_text(size=15,face="bold"), axis.text.y = element_text(size=15,face="bold")) +
                                 scale_color_manual(values=colors)
                                  

ggsave(file.path(plotpath,'REVAL_stability.png'),width = 6,height = 4)
gg.spaghetti_x
```

