---
title: "New_DP_finder"
author: "Veronica Mandelli"
date: "11/20/2022"
output: html_document
---
this script define peak of velocity and of acceleration of each trial's time series to later cut them into feedforward and feedback phases

# load libraries and set paths
```{r}
library(easypackages)
library(fpc)
library(readxl)
workdir = ""
libraries("ggplot2","dtw",'readr','patchwork', 'quantmod')
codepath = file.path(workdir,"code")
datapath = file.path(workdir,'data')
rawdatapath = file.path(workdir,"data","all_data_csv")
resultpath =  file.path(workdir,'result',"all_data")
cut_csv_path = file.path(workdir,"data",'all_data_csv_cutted')
plotpath = file.path(workdir,'plot')
plot_path_velocity_new = file.path(workdir,'plot','velocity_FINAL')
res_vel_path = file.path(workdir,'result','velocity_acceleration')
resultimetpath = file.path(workdir,'result')

# subject that have ABC and clusters
sub_info = read.csv("/Users/vmandelli/Google Drive/PHD/projects/motorNDAR/results/FINAL/dataset_bp_ASD_TD_clusters.csv")

# how many good trials for each subject (less that 10 will be excluded!)
sub_trials = read.csv2(file.path(rawdatapath,'subject_list.csv'))

# which trial is good/bad/missed
Trial_info =  read.csv2(file.path(rawdatapath,'TrialGoodBad.csv'))

# excel with timeframe to cut (start of the entire movement, end of the entire movement, end of reach, start of drop)
timetask <-read_excel(file.path(datapath,'timetask_alldata.xlsx'))
# how many subjects
length(unique(timetask$Soggetto))


##selcet only subjects that have also the MABC2 (So those that we have cluters labele)
sublist_def =read.csv(file.path(cut_csv_path,'sublist_def.csv'))
sublist_def = sublist_def$new_subID
outliers = c("1042",'1047','1088')
sublist_def=sublist_def[!sublist_def %in% outliers]
```

### reframe the dataset in a compehensible way
```{r}
# change subject names
sub_trials$Row <- gsub("RF_ASD_", "1",sub_trials$Row)
sub_trials$Row <- gsub("RF_asd_", "1",sub_trials$Row)

timetask$Soggetto <- gsub("asd", "1",timetask$Soggetto)
timetask$trial <- gsub("p", "",timetask$trial)
timetask$trial <-as.numeric(timetask$trial)

Trial_info$new_subID <- gsub("RF_asd_", "1",Trial_info$subID)
Trial_info$new_subID <- gsub("RF_ASD_", "1",Trial_info$new_subID)
```
#### uncomment here if you would like to run this analysis on TD 
```{r}
# # change subject names TD
# sub_trials$Row <- gsub("CT_", "2",sub_trials$Row)
# sub_trials$Row <- gsub("RFct_", "2",sub_trials$Row)
# 
# timetask$Soggetto <- gsub("ct", "2",timetask$Soggetto)
# timetask$trial <- gsub("p", "",timetask$trial)
# timetask$trial <- gsub("rova", "",timetask$trial)
# timetask$trial <-as.numeric(timetask$trial)
# 
# Trial_info$new_subID <- gsub("RFCT_", "2",Trial_info$subID)
# Trial_info$new_subID <- gsub("RFct_", "2",Trial_info$new_subID)
```

# select timetask (dataframe where time frame where manually inserted )
```{r}
col2use = c("Soggetto", "trial","LatoCastello","ManoPresa","Frame_Onset","Frame_Tocco","Frame_Onset2","Frame_End_Reaching2","FrameRilascio")
timetask = timetask[,col2use]
timetask = timetask[!duplicated(timetask),]
# select timeframes of subjects in 'sublist_def'
timetask_only10 = timetask[timetask$Soggetto %in% sublist_def,]

# work on Trial_info (table from matlab, that told us which trials has 3D data preprocessed)
Trial_info = Trial_info[Trial_info$new_subID %in% sublist_def,]
Trial_info = Trial_info[order(Trial_info$new_subID),]

```

# plot velocity path and compute the peaks
- velocity (max peaks velocity in the reaching and the in the dropping phases)
- deceleration peaks (the 1st peak after the velocity peaks, evetything that haoppen later is a sort of online correction (feedback processes))
it saves a plot of the velocity profiles of the wrist and the corresponding peaks.

```{r,error=FALSE,warning=FALSE}
sublist = sublist_def
library(dplyr)
plot_path_velocity_total= file.path(workdir,'plot','velocity_total')

exp_subj = c() 

# define the final subject list
sublist_ = sublist [!sublist %in% exp_subj]

# prepare a matrix to save the TimeFrames corresponding to the peaks and segmentation points
values2save = as.data.frame(matrix(ncol=8,nrow=0))
colnames(values2save) = c('sub_id','trial','p_vel_1','p_dec_1','end_1','start_2','p_vel_2','p_dec_2')

# start a for loop on all the subjects
for (k in c(1:length(sublist_))){
    # select the subject
    sub2use = sublist_[k]
    print(sub2use)
    # list the files for the subject (files altready cleaned (everything recorded before and after the real movemebt have been cutted out))
    names = paste(sub2use,"_*" ,sep="")
    trpresent = list.files(cut_csv_path,pattern=names)
    sub_timetask = subset(timetask_only10,timetask_only10$Soggetto == sub2use)
    
    
    #select the first 10 good trial for the subject from the csv
    t1 = read_csv(paste(cut_csv_path,'/',trpresent[1], sep=''))[,2:13] #2:13 are columns of body parts
    t2 = read_csv(paste(cut_csv_path,'/',trpresent[2], sep=''))[,2:13]
    t3 = read_csv(paste(cut_csv_path,'/',trpresent[3], sep=''))[,2:13]
    t4 = read_csv(paste(cut_csv_path,'/',trpresent[4], sep=''))[,2:13]
    t5 = read_csv(paste(cut_csv_path,'/',trpresent[5], sep=''))[,2:13]
    t6 = read_csv(paste(cut_csv_path,'/',trpresent[6], sep=''))[,2:13] 
    t7 = read_csv(paste(cut_csv_path,'/',trpresent[7], sep=''))[,2:13]
    t8 = read_csv(paste(cut_csv_path,'/',trpresent[8], sep=''))[,2:13]
    t9 = read_csv(paste(cut_csv_path,'/',trpresent[9], sep=''))[,2:13]
    t10 = read_csv(paste(cut_csv_path,'/',trpresent[10], sep=''))[,2:13]
    tlist= list(t1,t2,t3,t4,t5,t6,t7,t8,t9,t10)
    
    o=0
    trial2use=c()
    # select the  trails' number based on subject's specific trial list (not all the good trials are from 0 t 9, some start from 1, and some middle trials are missed) --> this is necessary to use the timetask to select the frames of: end of reaching and start of dropping, which are manually defined
    for(whichtrial in trpresent){
        o=o+1
        splitted = strsplit(whichtrial,split='_')
        trial2use[o]=splitted[[1]][3]
    }
    trial2use=as.numeric(trial2use)
    
    # select if the wrist (polso) used is SX or DX
    if (colnames(t1)[1]=="X_45metDx"){
        col2plot = c("X_PolsoMedDx","Y_PolsoMedDx","Z_PolsoMedDx")
    }else{col2plot = c("X_PolsoMedSn","Y_PolsoMedSn","Z_PolsoMedSn")}
    
    #initialize index and plot list
    plt_all=list()
    tr=0
    
    # loop on trials
    for (trial in tlist){
        tr = tr+1
        # initialize an empty matrix  to compute all the time series (velocity/acceleration)
        mat2save= as.data.frame(matrix(ncol=1,nrow=nrow(trial))) ### -1
        colnames(mat2save)= c('velocity')
        
        for (time_frame in 2: nrow(trial)){
            mat2save[time_frame,'timeframe']=time_frame
            # computing the istantaneous velocity 
            mat2save[time_frame,'velocity'] = sqrt((trial[time_frame,col2plot[1]]- trial[time_frame-1,col2plot[1]])^2 + (trial[time_frame,col2plot[2]]- trial[time_frame-1,col2plot[2]])^2+(trial[time_frame,col2plot[3]]- trial[time_frame-1,col2plot[3]])^2)/16.66
        }
        #add velocity = 0 in time frame 1
        mat2save[1,"velocity"]=0
        
        # find the acceleration ( the difference ad velocity at T - velocity at T-1)
        for (n in 2:length(mat2save$velocity)){
            mat2save[n,'acceleration']= (mat2save[n,'velocity']-mat2save[n-1,'velocity'])/16.66
        }
        
        # fill the 0 in the time frame -acceleration at 
        mat2save[1,"timeframe"]=1
        mat2save[1,"acceleration"]=0

        # find the peaks
        # velocity
        x= mat2save[,'velocity']
        peaks_vel = findPeaks(x, thresh=0)-1 # -1 because the function finds the timepoint after the peaks
        valley_vel = findValleys(x)-1
        # acceleration
        x= mat2save[,'acceleration']
        peaks_ac = findPeaks(x, thresh=0)-1
        # deceleration
        x= mat2save[,'acceleration']
        peaks_dec = findValleys(x)-1
        
        
        #  REACH find the peaks for the 1st movement 
        # define the TimeFrame of the touch (which represent the end of the reach)
        touch_frame = as.numeric(sub_timetask[sub_timetask$trial== as.numeric(trial2use[tr]),'Frame_Tocco']) - as.numeric(sub_timetask[sub_timetask$trial== as.numeric(trial2use[tr]),'Frame_Onset'])
        
        # peak velocity 
        peaks_vel_real = mat2save[mat2save$velocity == max(mat2save[mat2save$timeframe %in% peaks_vel[peaks_vel<as.numeric(touch_frame)],'velocity']),'timeframe']
        
        # Deceleration peak (real_dec) right after the peaks_vel_real
        real_dec_ = peaks_dec[peaks_dec>peaks_vel_real+1][1]   
        
        # DROP find the peaks for the 2nd movement
        # define the TimeFrame of the drop onset
        start_drp_frame =  as.numeric(sub_timetask[sub_timetask$trial== as.numeric(trial2use[tr]),'Frame_Onset2'])-as.numeric(sub_timetask[sub_timetask$trial== as.numeric(trial2use[tr]),'Frame_Onset']) 
        
        # peak velocity 
        peaks_vel_real_2 = mat2save[mat2save$velocity == max(mat2save[mat2save$timeframe %in% peaks_vel[peaks_vel>as.numeric(start_drp_frame)],'velocity']),'timeframe']
        
        # peak of decelertion (real_dec_2) right after the peaks_vel_real_2
        real_dec_2 = peaks_dec[peaks_dec>(peaks_vel_real_2+1) ][1]  
        
        #plot the time series 
        p = ggplot(mat2save, aes_string(x='timeframe', y='velocity')) +geom_line() +
            geom_vline(xintercept = touch_frame , color = "black", size=1.5) +
            geom_vline(xintercept = start_drp_frame , color = "black", size=1.5) +
            geom_vline(xintercept = peaks_vel_real, color = "blue", size=1.5) +
            geom_vline(xintercept = real_dec_ , linetype="dotted",color = "blue", size=1.5)+
            geom_vline(xintercept = peaks_vel_real_2, color = "red", size=1.5) +
            geom_vline(xintercept = real_dec_2 , linetype="dotted",color = "red", size=1.5) +
            xlab(as.character(tr))
        plt_all[[tr]] = p

        # eventually plot the acceleration profiles
        # p_ac = ggplot(mat2save, aes_string(x='timeframe', y='acceleration')) +geom_line() 
        
        #save the matrix of cutting timeframes
        values2save_ = as.data.frame(matrix(ncol=8,nrow=1))
        colnames(values2save_) = c('sub_id','trial','p_vel_1',
                                   'p_dec_1','end_1','start_2',
                                   'p_vel_2','p_dec_2')
        if (length(real_dec_)==0){
            real_dec_ = NaN
        }
        if (length(real_dec_2)==0){
            peaks_ac_real = NaN
        }
        values2save_[1,] = c(sub2use,trial2use[tr],peaks_vel_real,real_dec_,
                             touch_frame,start_drp_frame,
                             peaks_vel_real_2,real_dec_2)
        
        
        values2save = rbind(values2save,values2save_)
        # save each subject matrix
        write.csv(mat2save,file.path(res_vel_path,paste(sub2use,'_',tr,'vel_acc.csv')))
    }
    #bind plot of all trials to gether in a single page
    library(gridExtra)
    g = grid.arrange(plt_all[[1]],
                     plt_all[[2]],
                     plt_all[[3]],
                     plt_all[[4]],
                     plt_all[[5]],
                     plt_all[[6]],
                     plt_all[[7]],
                     plt_all[[8]],
                     plt_all[[9]],
                     plt_all[[10]],nrow=10
    )
    
    plt_subject=plt_all[[1]]+plt_all[[2]]+plt_all[[3]]+plt_all[[4]]+plt_all[[5]]+
        plt_all[[6]]+plt_all[[7]]+ plt_all[[8]]+ plt_all[[9]]+plt_all[[10]] + plot_layout(nrow = 10)
    
    #save the plot
    ggsave(file.path(plot_path_velocity_new,paste('sub_',sub2use,'_short.pdf',sep='')),g,width=3,height = 10)
}

#save total matrix for the subjetc
write_csv(values2save,file.path(resultimetpath,'times2cutDP.csv'))
```

## end